Simple SIR

### 2018-04-09 SimpleSIR.Rmd
Updated figure order and commented code

Deterministic and more simplistic stochastic SIR models for conspecific density vs. rec. rate combinations tested in spatially-explicit model

#Load libraries

```{r setup}
library(deSolve)
```

@@

freq_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/16512  # The change in S 
        dI = beta * S * I/16512 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

density_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/257  # The change in S 
        dI = beta * S * I/257 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

@@

## Stochastic SIR Model

```{r}
#' Function to randomly generate individuals and randomly assign 'nI' number of infected
#' individuals at t=0
#' @param n.initial- the number of desired individuals to start with
#' @param nI- the number of infected individuals to start with at t=0
#' @author Lauren White
make.inds <- function(n.initial, nI){
  id<-1:n.initial
  I<-sample(1:n.initial, nI)
  status<-rep("S", times=n.initial)
  status[I]<-"I"
  inds <- data.frame(id = id, status=status, stringsAsFactors=FALSE) 
  inds
}

@@

# tic=Sys.time()
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")

@@

```
Plot deterministic and stochastic SIR results (assuming equal mixing)
```{r}
par(mfrow=c(2, 3))
k<-8
dim<-2^k + 1
density<-0.25
n.initial25 <- round(density*dim^2) # inital population size

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

@@

##Make a merged dataset from all available .csv files
SIRdata_merged<- do.call("rbind", lapply(summaries, read.csv, header = TRUE))
#setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/")

@@

##Box plots
Box plots comparing density and recovery rates
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures Dec. 2017")
#sub_data<-merged_data[which(merged_data$max_I>1),] #successful outbreaks

@@

means <- aggregate(duration ~ density+ rec_rate, SIRdata_merged, mean)
#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
hh<- ggplot(data=SIRdata_merged, aes(x=as.factor(rec_rate), y=duration))
#hh<- hh+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "density", x = expression(Recovery~rate~(gamma)), y = "Duration", title="(B)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=16), title=element_text(face="bold", size=16))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=14))
hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           shape=18, size=3,show_guide = FALSE)
hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh
#dev.off()

@@

##Box plot
Compare homogenous mixing and spatial models
```{r}
SIRdata_merged$percep<-"Homogenous"
SIRdata_merged$p<-NA
SIRdata_merged$H<-NA
SIRdata_merged$beta1<-NA
SIRdata_merged$beta2<-NA
SIRdata_merged$beta3<-NA
SIRdata_merged$betas<-NA
SIRdata_merged$pbyH<-NA
SIRdata_merged$Hurst<-NA
SIRdata_merged$proportion<-NA
SIRdata_merged$lstructure<-NA

h_vs_s <-rbind(sub_data, SIRdata_merged)
hh <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = max_prevalence, fill = percep))
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Maximum prevalence", title="(A)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=16), title=element_text(face="bold", size=16))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=14))
#hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh

ii <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = duration, fill = percep))
ii<-ii+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
ii<- ii+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Duration", title="(B)")
ii<- ii+ theme(axis.title = element_text(face="bold", size=16), title=element_text(face="bold", size=16))
ii<- ii + theme(axis.text.x = element_text(vjust=0.5, size=16))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=14))
#ii<- ii + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#ii<-ii + geom_text(data = means , aes(label = duration, y = duration + 0.08))
ii

setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures Dec. 2017")
tiff("Compare_homog_spatial.tif", height = 8, width = 14, units = "in", compression = "lzw", res = 300)
multiplot(hh,ii, cols=2)
dev.off()
```

>>>>

Deterministic and more simplistic stochastic SIR models for conspecific density vs. rec. rate combinations tested in spatially-explicit model. Assume homogenous mixing and no landscape structure.

#Load libraries
Load `deSolve` library for evaluating ordinary differntial equations (ODEs).
```{r setup}
library(deSolve)
library(ggplot2)
library(cowplot)
```

@@

#frequency dependent transmission
freq_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/16512  # The change in S 
        dI = beta * S * I/16512 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

#density dependent transmission function
density_sir <- function(time, state, parameters) {
    with(as.list(c(state, parameters)), {
        dS = -beta * S * I/257  # The change in S 
        dI = beta * S * I/257 - gamma * I
        dR = gamma * I
        return(list(c(dS, dI, dR))) 
    })
}

@@

## Stochastic SIR Model 
Develop functions to run equivalent stochastic SIR model that assumes density-dependent transmission, homogenous mixing, and operates using a Reed Frost model.
```{r}
#' Function to randomly generate individuals and randomly assign 'nI' number of infected
#' individuals at t=0
#' @param n.initial- the number of desired individuals to start with
#' @param nI- the number of infected individuals to start with at t=0
#' @author Lauren White
make.inds <- function(n.initial, nI){
  id<-1:n.initial
  I<-sample(1:n.initial, nI)
  status<-rep("S", times=n.initial)
  status[I]<-"I"
  inds <- data.frame(id = id, status=status, stringsAsFactors=FALSE) 
  inds
}

@@

##Run stochastic simulations
Run simplified SIR model for six combinations of conspecific density and recovery rate tested in spatially explicit model.

```{r}
# tic=Sys.time()
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")

@@

##Plot deterministic and stochastic SIR results (assuming equal mixing)
Compare/verify output from deterministic ODE and stochastic SIR models
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/SimpleSIR")
#set wd to location of stochastic SIR simulation output files

par(mfrow=c(2, 3))
k<-8
dim<-2^k + 1
density<-0.25
n.initial25 <- round(density*dim^2) # inital population size

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

@@

stoch<-read.csv("k8dens0.25inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.1_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.1))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.2_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.2))
head(out)

@@

stoch<-read.csv("k8dens0.5inf0.2rec0.4_infected.csv")
out <- ode(y = init, func = density_sir, times = times, parms = c(beta = 0.2, gamma = 0.4))
head(out)

@@

##Make a merged dataset from all available .csv files
SIRdata_merged<- do.call("rbind", lapply(summaries, read.csv, header = TRUE))
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/")
write.csv(SIRdata_merged, file="SIRdata_merged.csv")

@@

##Box plots
Box plots comparing density and recovery rates for homogenous mixing SIR model
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures Dec. 2017")
#sub_data<-merged_data[which(merged_data$max_I>1),] #successful outbreaks

@@

means <- aggregate(duration ~ density+ rec_rate, SIRdata_merged, mean)
#tiff("boxplot_dur.tiff", height = 8, width = 12, units = "in", compression = "lzw", res = 300)
hh<- ggplot(data=SIRdata_merged, aes(x=as.factor(rec_rate), y=duration))
#hh<- hh+ geom_bar(stat="identity", position=position_dodge(), colour="black")+ facet_grid(Rec.prob. ~ percep, scales= "free_x", labeller=label_both)
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "density", x = expression(Recovery~rate~(gamma)), y = "Duration", title="(B)")
hh<- hh+ theme(axis.title = element_text(face="bold", size=16), title=element_text(face="bold", size=16))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=16))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=16)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=14))
hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           shape=18, size=3,show.legend = FALSE)
hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh
#dev.off()

@@

##Fig. S6
Compare homogenous mixing and spatial models in box plot form.
```{r}
setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/")
SIRdata_merged<-read.csv("SIRdata_merged.csv")
SIRdata_merged<-SIRdata_merged[,-1]

SIRdata_merged$percep<-"H"
SIRdata_merged$p<-NA
SIRdata_merged$H<-NA
SIRdata_merged$beta1<-NA
SIRdata_merged$beta2<-NA
SIRdata_merged$beta3<-NA
SIRdata_merged$betas<-NA
SIRdata_merged$pbyH<-NA
SIRdata_merged$Hurst<-NA
SIRdata_merged$proportion<-NA
SIRdata_merged$lstructure<-NA

h_vs_s <-rbind(merged_data, SIRdata_merged)
hh <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = max_prevalence, fill = percep))
hh<-hh+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
hh<- hh+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Maximum prevalence")#, title="(A)")
hh<- hh+ theme(axis.title = element_text(size=6), plot.title=element_text(size=7))
hh<- hh + theme(axis.text.x = element_text(vjust=0.5, size=6))
hh<- hh + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
hh<- hh + theme(strip.text = element_text(vjust=0.5, size=6))
hh<- hh + theme(legend.position="none")
hh<- hh + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
hh<- hh +theme(plot.margin=unit(c(0.1,0.1,0,0.1),"cm"))
#hh<- hh + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#hh<-hh + geom_text(data = means , aes(label = duration, y = duration + 0.08))
hh

ii <- ggplot(h_vs_s, aes(x = as.factor(rec_rate), y = duration, fill = percep))
ii<-ii+ geom_boxplot()+ facet_grid(. ~ density, scales= "free_x", labeller=label_both)
ii<- ii+ labs(fill= "Perceptual \ndistance", x = expression(Recovery~rate~(gamma)), y = "Duration") #, title="(B)")
ii<- ii+ theme(axis.title = element_text(size=6), plot.title=element_text(size=7))
ii<- ii + theme(axis.text.x = element_text(vjust=0.5, size=6))
ii<- ii + theme(axis.text.y = element_text(vjust=0.5, size=6)) 
ii<- ii + theme(strip.text = element_text(vjust=0.5, size=6))
ii<- ii + theme(legend.position="bottom")
ii<- ii + theme(legend.text = element_text(vjust=0.5, size=6), legend.title=element_text(vjust=0.5, size=7))
ii<- ii +theme(plot.margin=unit(c(-0.1,0.1,-0.5,0.1),"cm"))
#ii<- ii + stat_summary(fun.y=mean, colour="darkred", geom="point", 
                           #shape=18, size=3,show_guide = FALSE)
#ii<-ii + geom_text(data = means , aes(label = duration, y = duration + 0.08))
ii

#install.packages("ggpubr")
#library("ggpubr")
#ggarrange(hh, ii, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")

library(cowplot)
prow <- plot_grid( hh + theme(legend.position="none"),
            ii + theme(legend.position="none"),
           align = 'vh',
           labels = c("(A)", "(B)"),
           hjust = -1,
           nrow = 2, label_size=7
           )

# extract the legend from one of the plots
# (clearly the whole thing only makes sense if all plots
# have the same legend, so we can arbitrarily pick one.)
legend_b <- get_legend(ii + theme(legend.position="bottom"))

# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
p <- plot_grid( prow, legend_b, ncol = 1, rel_heights = c(1, .07))
p

setwd("C:/Users/law2y/OneDrive/R Code/LandscapeCh/Figures April 2018")
tiff("FigureS6.tiff", height = 17, width = 17,8, units = "cm", compression = "lzw", res = 1200)
p
#multiplot(hh,ii, cols=1)
dev.off()
```
